\documentclass[a4paper,10pt]{article}


%opening
\title{Math650 Homework 3.2}
\author{Yu Huang}

\begin{document}

\maketitle

\begin{abstract}
Speed Limits and Traffic Fatalities
\end{abstract}

\section{Introduction}
Here we investigate the question whether the increase of speed limits caused the traffic fatalities to increase.

\section{Materials and Methods}
Question Chap2 No23, Data from its CDROM and R software.


\section{Results}
Sample 1 is percentage change data of states which increased speed limits. Sample 2 is data of states which didn't change speed limits.

Result Table
\begin{tabular}{|r|r|}
\hline
mean of sample 1 & 13.75312\\
mean of sample 2 & -8.563158\\
mean difference & 22.31628\\
sd of sample 1& 21.33285\\
sd of sample 2& 31.00085\\
degrees of freedom of pooled sd & 49\\
pooled sd & 25.31707\\
standard error for the difference & 7.33241\\
t statistic & 3.043513\\
p-value of t statistic & 0.001876831\\
\hline
\end{tabular}

All R codes are appended(\ref{appendix}).

\section{Conclusion and Discussion}
Here there's neither random selection nor random allocation of units to groups. We can't directly say there's a strong causal relationship between speed limits and traffic fatalities. However, based on the tiny p-value(0.001876831), the actual data differs substantially from the expected outcome under the model which assumes no increase of traffic fatalities after increasing speed limits.

\section{Appendix}
\label{appendix}
function \emph{t\_test\_func} is same as in math650\_hw3\_1.
\begin{verbatim}
#chap 2, No23
data = read.csv("/usr/local/doc/statistical_sleuth/ASCII/ex0223.csv")
sample_data1 = data[data$INCREASe=="Yes",]$FATALITIESCHANGE
sample_data2 = data[data$INCREASe=="No",]$FATALITIESCHANGE
t_test_func(sample_data1, sample_data2)
\end{verbatim}

Output is this (disregarding those lines with df=12):
\begin{verbatim}
mean_f: 13.75312
mean_r: -8.563158
mean difference: 22.31628
sd_f: 21.33285
sd_r: 31.00085
degrees_of_freedom_of_pooled_sd: 49
pooled_sd: 25.31707
standard_error_for_the_difference: 7.33241
percentile_97_5th, df=12: 2.179
conf_interv_of_difference_of_mu_lower, df=12: 6.338962
conf_interv_of_difference_of_mu_upper, df=12: 38.2936
t_stat: 3.043513
percentile_97_5th: 2.009575
conf_interv_of_difference_of_mu_lower: 7.581254
conf_interv_of_difference_of_mu_upper: 37.05131
p-value of t_stat 0.001876831
\end{verbatim}
\end{document}
